---
title: "Cookbook"
output:
  html_document:  
    toc: true  
    toc_float: true 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(extrafont)
```


These examples all use the "diamonds" dataset from the ggplot package.

## Font

The `extrafont` package enables the use of custom fonts in the plots. To install a new font, download the TrueType font file, and follow the instructions here.

## Scatter Plot

```{r Scatter, echo=FALSE, message=FALSE, warning=FALSE}
# Set colour scheme
se_colour <- c("#E79AB5", "#B8345C", "#8CCDD6", "#43A8B6", "#00818F", "#C1DC9A", "#7ABB40", "#5D4086") # colours for use in plot

# Scatter plot
d <- diamonds
plot.point <- ggplot(data = d)+ 
  geom_point(aes(x = carat, y = price, color = cut)) + 
  theme(text = element_text(family = "Norwester", color = "#5F5F5F"), 
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5), 
        panel.background  =  element_blank() ) + 
  labs( x= "Carat", y = "Price", title = "Price by Carat", fill = "Cut") +
  scale_color_manual(values= se_colour) 
plot.point
```

Depending on your data, other additional changes may be required to "neaten" the plot. For example, we can reformat the y axis labels to display them as currency:

```{r Scatter2, echo=TRUE}

# Scatter plot
d <- diamonds
plot.point.v1 <- ggplot(data = d)+ 
  geom_point(aes(x = carat, y = price, color = cut)) + 
  theme(text = element_text(family = "Norwester", color = "#5F5F5F"), 
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5), 
        panel.background  =  element_blank() ) + 
  labs( x= "Carat", y = "Price", title = "Price by Carat", fill = "Cut") +
  scale_color_manual(values= se_colour) +
  scale_y_continuous(labels = scales::dollar_format(prefix = "£")) # reformat scales to currency
plot.point.v1

```
## Boxplot

```{r Boxplot, echo=TRUE}

plot.box <- ggplot(data = d)+ 
  geom_boxplot(aes(x = carat, y = price, fill = cut)) + 
  theme(text = element_text(family = "Norwester", color = "#5F5F5F"), 
  plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5),
        panel.grid = element_blank()) + 
  labs( x= "Carat", y = "Price", title = "Price by Carat", fill = "Cut") +
  scale_fill_manual(values=se_colour) +
  scale_y_continuous(labels = scales::dollar_format(prefix = "£")) # reformat scales to currency
plot.box


```

## Histogram

```{r}
plot.hist <- ggplot(data = d)+ 
  geom_histogram(aes(x = price, fill = cut), binwidth = 200) + 
  theme(text = element_text(family = "Norwester", color = "#5F5F5F"), 
  plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5),
        panel.grid = element_blank()) + 
  labs( x= "Price", y = "Number of Diamonds", title = "Price by Cut", fill = "Cut") +
  scale_fill_manual(values=se_colour)+
  scale_x_continuous(labels = scales::dollar_format(prefix = "£")) # reformat scales to currency
plot.hist

```

