---
title: "Leaflet"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(jsonlite)
library(dplyr)
library(tidyr)
library(extrafont)
library(leaflet)

#access url
base2 <- "http://api.letsride.co.uk/public/v1/rides"
openactive <- GET(base2) 


#from unicode to text
this.raw.content <- rawToChar(openactive$content)
nchar(this.raw.content) #size
substr(this.raw.content, 1, 100) #first 100 characters

#parse to JSON
this.content <- fromJSON(this.raw.content)

#extract geo information 
newdata <- this.content$items
newdata1 <- newdata$data
newdata2 <- newdata1$location
newdata3 <- newdata2$geo

#bind dataframes
newdata4 <- cbind(newdata1, newdata3)
newdata4 <- newdata4[c(2, 14, 27, 23, 35, 36)]

#rename Gender categories
newdata4$genderRestriction[newdata4$genderRestriction == "https://openactive.io/FemaleOnly"] <- "Women-only Session"
newdata4$genderRestriction[is.na(newdata4$genderRestriction)] <- "Mixed-Session"

mses <- subset(newdata4, genderRestriction == "Mixed-Session") 
wses <- subset(newdata4, genderRestriction == "Women-only Session")

```

The *Leaflet* package creates a map widget; you can add layers such as tiles, markers, and polygons to build up the map. There is already a fantastic guide for using [Leaflet](https://rstudio.github.io/leaflet/), but here is a quick overview on how to create a simple map, using data from British Cycling via the [Open Active API]("http://api.letsride.co.uk/public/v1/rides"). 

<div style="text-align:left"><span style="color:black; font-size:1em;">Step 1: Load the library</span></div>

All we need to create the map is a dataset with longitude and latitude (which we have from the Open Active API) and the *Leaflet* package.

```{r library, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(leaflet)
```

<div style="text-align:left"><span style="color:black; font-size:1em;">Step 2: Get the data</span></div>

Here's a table of my British Cycling session information:

```{r table, echo=FALSE}
head(newdata4[,c(1,3,5,6)])
```



<div style="text-align:left"><span style="color:black; font-size:1em;">Step 3: Create your map using Leaflet</span></div>
Below is a map widget with Open Street Map tiles (the default) and circle markers of different colours to differentiate between mixed sessions, and women-only sessions:

```{r pressure, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# use leaflet to map
leaflet() %>% addTiles() %>%
  addCircleMarkers(data= mses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction), color = "#43A8B6")  %>%
  addCircleMarkers(data= wses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction),  color = "#5D4086")
```

It is also possible to cluster these markers by using the *clusterOptions* argument:

```{r}
leaflet() %>% addTiles() %>%
  addCircleMarkers(data= mses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction), color = "#43A8B6", clusterOptions = markerClusterOptions())  %>%
  addCircleMarkers(data= wses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction),  color = "#5D4086", clusterOptions = markerClusterOptions())

```

