---
title: "Leaflet Tutorial"
output:
  html_document:  
    toc: true  
    toc_float: true 
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(jsonlite)
library(dplyr)
library(tidyr)
library(extrafont)
library(leaflet)
library(leaflet.esri)
library(kableExtra)

#access url
base2 <- "http://api.letsride.co.uk/public/v1/rides"
openactive <- GET(base2) 

#from unicode to text
this.raw.content <- rawToChar(openactive$content)

#parse to JSON
this.content <- fromJSON(this.raw.content)

#extract geo information 
newdata <- this.content$items
newdata1 <- newdata$data
newdata2 <- newdata1$location
newdata3 <- newdata2$geo

#bind dataframes
newdata4 <- cbind(newdata1, newdata3)


#rename Gender categories
newdata4$genderRestriction[newdata4$genderRestriction == "https://openactive.io/FemaleOnly"] <- "Women-only Session"
newdata4$genderRestriction[is.na(newdata4$genderRestriction)] <- "Mixed-Session"

mses <- subset(newdata4, genderRestriction == "Mixed-Session") 
wses <- subset(newdata4, genderRestriction == "Women-only Session")

df <- read.csv('input/All_Lottery_18_20_latlon.csv', header = TRUE)

```

# About Leaflet
The `Leaflet` package creates a map widget; you can add layers such as tiles, markers, and polygons to build up the map. There is already a fantastic guide for using [Leaflet](https://rstudio.github.io/leaflet/), but here is a quick overview on how to create a simple map, using data from British Cycling via the [Open Active API]("http://api.letsride.co.uk/public/v1/rides"). 

# Load the Library

All we need to create the map is a dataset with longitude and latitude (which we have from the Open Active API) and the `Leaflet` package.

```{r library, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(leaflet)
```

# Get the data

Here's a table of my British Cycling session information:

```{r table, echo=FALSE, fig.align='center'}
kable(newdata4[1,c(27,35,36)]) %>%
  kable_styling(bootstrap_options = "bordered",
                full_width = FALSE) 
```


# Create your map

Below is a map widget with Open Street Map tiles (the default) and circle markers of different colours to differentiate between mixed sessions, and women-only sessions:

```{r pressure, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.align='center'}
# use leaflet to map
leaflet() %>% addTiles() %>%
  addCircleMarkers(data= mses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction), color = "#43A8B6")  %>%
  addCircleMarkers(data= wses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction),  color = "#5D4086")
```

It is also possible to cluster these markers by using the `clusterOptions` argument:

```{r leaf, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.align='center'}
leaflet() %>% addTiles() %>%
  addCircleMarkers(data= mses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction), color = "#43A8B6", clusterOptions = markerClusterOptions())  %>%
  addCircleMarkers(data= wses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction),  color = "#5D4086", clusterOptions = markerClusterOptions())

```

# Additional features
There are numerous other features available in the `Leaflet` package, such as:
* Changing the basemap tile 
* Adding popups 
* Adding a legend
* Adding Esri feature layers using the `leaflet.esri` package

The Esri feature layers enables a user to overlay the dataset with additonal geographic data, such as green space, or Index of Multiple Deprivation. The below example uses Sport England Lottery Awards data:
```{r esri, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.align='center'}

df_19 <- df %>%
  filter(Year == "2019/20")

pal <- colorFactor(palette = c('#fff7f3','#fde0dd','#fcc5c0','#fa9fb5','#f768a1','#dd3497','#ae017e','#7a0177','#49006a', '#49006a'), domain = df$Index.of.Multiple.Deprivation.Decile)

m <- leaflet()%>%
  addTiles(group = "OSM")%>%
  addProviderTiles(providers$Esri.WorldGrayCanvas, group = "Esri Gray") %>%
  addCircleMarkers(data= df_19, ~Longitude, ~Latitude, 
                   popup = paste("Year:", df_19$Year, "<br>",
                                 "Amount:", "£", formatC(as.numeric(df$Amount.granted), big.mark = ",", digits = 10),  "<br>",
                                 "LSOA Name:", df_19$`LSOA.Name`), 
                   color = ~pal(df_19$Index.of.Multiple.Deprivation.Decile),
                   fillOpacity = ~pal(Index.of.Multiple.Deprivation.Decile),
                   radius = ~sqrt(Amount.granted)* 0.005) %>%
  addLegend("bottomright", pal = pal, values = df_19$Index.of.Multiple.Deprivation.Decile, title = "IMD",
            opacity = 1)%>%
  addEsriFeatureLayer(url = "https://services3.arcgis.com/ivmBBrHfQfDnDf8Q/ArcGIS/rest/services/Indices_of_Multiple_Deprivation_(IMD)_2019/FeatureServer/0",
                     weight = 1, color = "grey",
                     labelProperty = paste("LSOA01NM"), 
                     group = "IMD 2019 Layer",
                     labelOptions = labelOptions(textsize = "12px"),
                     popupProperty = JS("function(feature) { return L.Util.template(\"<h3>IMD Decile: {IMDDec0}</h3><hr /><p>This LSOA is {LSOA01NM} and the IMD Rank is {IMDRank0}. The total population is {TotPop} (excluding pensioners)\", feature.properties);}")) %>%
  
   addEsriFeatureLayer(url = "https://ons-inspire.esriuk.com/arcgis/rest/services/Administrative_Boundaries/Local_Authority_Districts_December_2017_Boundaries_WGS84/MapServer/3",
                     weight = 1, color = "grey",
                     labelProperty = paste("lad17nm"), 
                     group = "LA 2017 Layer",
                     labelOptions = labelOptions(textsize = "12px")) %>%
addEsriFeatureLayer(url = "https://services5.arcgis.com/piNGxor37zqmvRAm/ArcGIS/rest/services/OS_Open_Greenspace/FeatureServer/1",
                     weight = 1, color = "grey",
                     labelProperty = paste("function_"), 
                     group = "Green Space",
                     labelOptions = labelOptions(textsize = "12px"))%>%
  addLayersControl(
    baseGroups = c("OSM", "Esri Gray"),
    overlayGroups = c("LA 2017 Layer", "IMD 2019 Layer", "Green Space"),
    options = layersControlOptions(collapsed = FALSE)
  )

m %>% hideGroup(c ("LA 2017 Layer", "IMD 2019 Layer", "Green Space"))



```
