---
title: "Leaflet Tutorial"
output:
  html_document:  
    toc: true  
    toc_float: true 
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(jsonlite)
library(dplyr)
library(tidyr)
library(extrafont)
library(leaflet)
library(kableExtra)

#access url
base2 <- "http://api.letsride.co.uk/public/v1/rides"
openactive <- GET(base2) 

#from unicode to text
this.raw.content <- rawToChar(openactive$content)

#parse to JSON
this.content <- fromJSON(this.raw.content)

#extract geo information 
newdata <- this.content$items
newdata1 <- newdata$data
newdata2 <- newdata1$location
newdata3 <- newdata2$geo

#bind dataframes
newdata4 <- cbind(newdata1, newdata3)


#rename Gender categories
newdata4$genderRestriction[newdata4$genderRestriction == "https://openactive.io/FemaleOnly"] <- "Women-only Session"
newdata4$genderRestriction[is.na(newdata4$genderRestriction)] <- "Mixed-Session"

mses <- subset(newdata4, genderRestriction == "Mixed-Session") 
wses <- subset(newdata4, genderRestriction == "Women-only Session")

```

# About Leaflet
The `Leaflet` package creates a map widget; you can add layers such as tiles, markers, and polygons to build up the map. There is already a fantastic guide for using [Leaflet](https://rstudio.github.io/leaflet/), but here is a quick overview on how to create a simple map, using data from British Cycling via the [Open Active API]("http://api.letsride.co.uk/public/v1/rides"). 

# Load the Library

All we need to create the map is a dataset with longitude and latitude (which we have from the Open Active API) and the `Leaflet` package.

```{r library, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(leaflet)
```

# Get the data

Here's a table of my British Cycling session information:

```{r table, echo=FALSE, fig.align='center'}
kable(newdata4[1,c(27,35,36)]) %>%
  kable_styling(bootstrap_options = "bordered",
                full_width = FALSE) 
```


# Create your map

Below is a map widget with Open Street Map tiles (the default) and circle markers of different colours to differentiate between mixed sessions, and women-only sessions:

```{r pressure, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.align='center'}
# use leaflet to map
leaflet() %>% addTiles() %>%
  addCircleMarkers(data= mses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction), color = "#43A8B6")  %>%
  addCircleMarkers(data= wses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction),  color = "#5D4086")
```

It is also possible to cluster these markers by using the `clusterOptions` argument:

```{r leaf, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.align='center'}
leaflet() %>% addTiles() %>%
  addCircleMarkers(data= mses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction), color = "#43A8B6", clusterOptions = markerClusterOptions())  %>%
  addCircleMarkers(data= wses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction),  color = "#5D4086", clusterOptions = markerClusterOptions())

```

